---
title: "CalculatePi1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

## Introduction

I have been making a lot of plots of heatmaps to summarise colocalization rate between different types of QTLs. Here I will make some analogous plots with the pi1 statistic. I already made a file that take all pairs of traits that i attempted to coloc, and collects the P value of trait2 (trait.y) for the top snp in trait.x. I will use 

```{r}
library(tidyverse)
library(qvalue)
library(gplots)
library(viridis)
library(data.table)

```

Read in data

```{r}

dat <- fread("../code/scratch/PairwiseColocStats.AddedAscertainmentP.txt.gz", sep='\t')

RecodeDat <- read_tsv("../data/Phenotypes_recode_for_Plotting.txt")

RecodeVec <- RecodeDat %>%
  select(PC, ShorterAlias) %>%
  deframe()

RecodeIncludePCs <- RecodeDat %>%
  filter(Include) %>%
  pull(PC)



RNASeqPhenotypes <- c("MetabolicLabelled.30min", "MetabolicLabelled.60min", "Expression.Splicing.Subset_YRI", "chRNA.Expression.Splicing", "H3K36ME3")



```

First some sanity checks:

what is the distribution of p values in phenotype.y, for top snps for phenotype.x

```{r}
hist(dat$trait.x.p.in.y)
```

...and what for colocalized trait pairs versus non colocalied trait pairs:

```{r}
dat %>%
  ggplot(aes(x=trait.x.p.in.y, fill=IsColocalizedPair)) +
  geom_histogram() +
  facet_wrap(~IsColocalizedPair, scales = "free_y") +
  labs(x="Pvalue", y="count", title="Trait.y P values at the top SNP for trait.x") +
  theme_classic()

```

Now about the one phenotype to many problem... first let's see the distribution of the number of phenotype y's compared to each phenotype x.

```{r}
dat %>%
  count(P1, PC1, PC2) %>%
  ggplot(aes(x=n)) +
  stat_ecdf() +
  coord_cartesian(xlim=c(0,25)) +
  labs(title="Most traits only attempted to colocalize one or a few\nof other traits from each class",
       x="Number of traits in a class (ie splicing)\n attempted to colocalize with each trait (ie polyA expression of a gene)",
       y="ecdf") +
  theme_bw()
```


Cactch errors and create pi1 heatmap without accounting for the one to many phenotypes problem...

```{r}

testFunction <- function (dat.in) {
  return(tryCatch(1-qvalue(dat.in$Pvals.For.Pi1)$pi0, error=function(e) NULL))
}

dat.split <- dat %>%
  filter(PC1 %in% RecodeIncludePCs) %>%
  filter(PC2 %in% RecodeIncludePCs) %>%
  mutate(PC1 = recode(PC1, !!!RecodeVec)) %>%
  mutate(PC2 = recode(PC2, !!!RecodeVec)) %>%
  select(PC1, PC2, Pvals.For.Pi1=trait.x.p.in.y) %>%
  filter(!PC1==PC2) %>%
  split(paste(.$PC1, .$PC2, sep = ";"))

dat.pi1 <- lapply(dat.split, testFunction) %>%
  unlist() %>%
  data.frame(pi1=.) %>%
  rownames_to_column("PC1_PC2") %>%
  separate(PC1_PC2, into=c("PC1", "PC2"), sep=';')

pi.heatmap <- ggplot(dat.pi1, aes(x=PC1, y=PC2, fill=pi1)) +
  geom_raster() +
  geom_text(aes(label=signif(pi1*100, 2)), color="blue") +
  scale_fill_viridis(option="B", direction = 1, limits=c(0,1)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
pi.heatmap  

```

now accounting for that problem...

```{r, eval=F}
### This code raises issues because qvalue throws errors when distribution has too much pi1
dat.split <- dat %>%
  filter(PC1 %in% RecodeIncludePCs) %>%
  filter(PC2 %in% RecodeIncludePCs) %>%
  mutate(PC1 = recode(PC1, !!!RecodeVec)) %>%
  mutate(PC2 = recode(PC2, !!!RecodeVec)) %>%
  mutate(runif = runif(n())) %>%
  mutate(runif2 = runif(n())) %>%
  group_by(PC1, P1, PC2) %>%
  mutate(test.stat.obs = -log10(min(trait.x.p.in.y)),
         test.stat.null = -log10(min(runif))) %>%
  ungroup() %>%
  group_by(PC1, PC2) %>%
  mutate(Pvals.For.Pi1 = empPvals(test.stat.obs, test.stat.null)) %>%
  ungroup() %>%
  add_count(PC1, PC2) %>%
  filter(n >= 100) %>%
  # filter((PC1 %in% RNASeqPhenotypes) & (PC2 %in% RNASeqPhenotypes)) %>%
  select(PC1, PC2, Pvals.For.Pi1) %>%
  filter(!PC1==PC2) %>%
  group_by(PC1, PC2) %>%
  summarise(pi1=case_when(
    max(Pvals.For.Pi1) < 0.05 ~ 1,
    TRUE ~ 1-qvalue(Pvals.For.Pi1)$pi0
    ))


pi.heatmap <- ggplot(dat.pi1, aes(x=PC1, y=PC2, fill=pi1)) +
  geom_raster() +
  geom_text(aes(label=signif(pi1*100, 2)), color="blue") +
  scale_fill_viridis(option="B", direction = 1, limits=c(0,1)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
pi.heatmap  

```

Now use that approach to catch errors with qvalue, while accounting for the one to many phenotypes problem.

```{r}
### This code works

dat.split <- dat %>%
  filter(PC1 %in% RecodeIncludePCs) %>%
  filter(PC2 %in% RecodeIncludePCs) %>%
  mutate(PC1 = recode(PC1, !!!RecodeVec)) %>%
  mutate(PC2 = recode(PC2, !!!RecodeVec)) %>%
  mutate(runif = runif(n())) %>%
  mutate(runif2 = runif(n())) %>%
  group_by(PC1, P1, PC2) %>%
  mutate(test.stat.obs = -log10(min(trait.x.p.in.y)),
         test.stat.null = -log10(min(runif))) %>%
  ungroup() %>%
  group_by(PC1, PC2) %>%
  mutate(Pvals.For.Pi1 = empPvals(test.stat.obs, test.stat.null)) %>%
  ungroup() %>%
  select(PC1, PC2, Pvals.For.Pi1) %>%
  filter(!PC1==PC2) %>%
  split(paste(.$PC1, .$PC2, sep = ";"))

dat.pi1 <- lapply(dat.split, testFunction) %>%
  unlist() %>%
  data.frame(pi1=.) %>%
  rownames_to_column("PC1_PC2") %>%
  separate(PC1_PC2, into=c("PC1", "PC2"), sep=';')

pi.heatmap <- ggplot(dat.pi1, aes(x=PC1, y=PC2, fill=pi1)) +
  geom_raster() +
  geom_text(aes(label=signif(pi1*100, 2)), color="blue") +
  scale_fill_viridis(option="B", direction = 1, limits=c(0,1)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Discovery QTL phenotype", y="Phenotype assessed for overlap")
pi.heatmap  

```

I think the above heatmaps are questionable because the size of the empirical null distribution is really small in most cases (ie 1 or 2), and I'm not sure how well empPvals behaves or introduces bias with this small null distribution. I think I will pre-calculate the null distribution for all samples sizes less than say 100.

```{r}

MaxSampleSizeToCreateANull <- 100
NSamplesToEstimateDistribution <- 10000
NullSimulatedTestStats <- matrix(nrow=MaxSampleSizeToCreateANull, ncol=NSamplesToEstimateDistribution)
rownames(NullSimulatedTestStats) <- paste0("runif_samplesize", 1:MaxSampleSizeToCreateANull)
colnames(NullSimulatedTestStats) <- paste0("Sample_", 1:NSamplesToEstimateDistribution)


for (i in 1:MaxSampleSizeToCreateANull){
  SampleSizeFromUniform <- i
  SampledDat <- matrix(runif(SampleSizeFromUniform*NSamplesToEstimateDistribution), nrow=NSamplesToEstimateDistribution)
  SampleDatNullTestStatistics <- -log10(apply(SampledDat, 1, min))
  NullSimulatedTestStats[i,] <- SampleDatNullTestStatistics
}

NullSimulatedTestStats %>%
  as.data.frame() %>%
  rownames_to_column("runif_samplesize") %>%
  slice(1:10) %>%
  gather(key="Sample", value="value", -runif_samplesize) %>%
  ggplot(aes(x=value)) +
  geom_histogram() +
  facet_wrap(~runif_samplesize, scales="free")

NullSimulatedTestStats %>%
  as.data.frame() %>%
  rownames_to_column("runif_samplesize") %>%
  slice(1:20) %>%
  # mutate(runif_samplesize = as.numeric(str_remove(runif_samplesize, "runif_samplesize"))) %>%
  gather(key="Sample", value="value", -runif_samplesize) %>%
  ggplot(aes(x=value, color=runif_samplesize)) +
  geom_density() +
  theme_bw()

```

Now let's remake the heatmaps

```{r, eval=F}

NullSimulatedTestStats

dat.split <- dat %>%
  # head(100000) %>%
  filter(PC1 %in% RecodeIncludePCs) %>%
  filter(PC2 %in% RecodeIncludePCs) %>%
  mutate(PC1 = recode(PC1, !!!RecodeVec)) %>%
  mutate(PC2 = recode(PC2, !!!RecodeVec)) %>%
  mutate(runif = runif(n())) %>%
  mutate(runif2 = runif(n())) %>%
  group_by(PC1, P1, PC2) %>%
  mutate(test.stat.obs = -log10(min(trait.x.p.in.y))) %>%
  ungroup() %>%
  add_count(PC1, P1, PC2) %>%
  filter(n<=100) %>%
  group_by(PC1, PC2) %>%
  rowwise() %>%
  mutate(Pvals.For.Pi1 = empPvals(test.stat.obs, NullSimulatedTestStats[n,])) %>%
  ungroup() %>%
  select(PC1, PC2, Pvals.For.Pi1) %>%
  filter(!PC1==PC2) %>%
  split(paste(.$PC1, .$PC2, sep = ";"))

dat.pi1 <- lapply(dat.split, testFunction) %>%
  unlist() %>%
  data.frame(pi1=.) %>%
  rownames_to_column("PC1_PC2") %>%
  separate(PC1_PC2, into=c("PC1", "PC2"), sep=';')

pi.heatmap <- ggplot(dat.pi1, aes(x=PC1, y=PC2, fill=pi1)) +
  geom_raster() +
  geom_text(aes(label=signif(pi1*100, 2)), color="blue") +
  scale_fill_viridis(option="B", direction = 1, limits=c(0,1)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Discovery QTL phenotype", y="Phenotype assessed for overlap")
pi.heatmap  
```


ooops i realized now that all this doesn't make sense since i am only calculating pi1 from trait pairs i attempted to colocalize, meaning they are definitely a QTL for something even in the Trait.y, which will upwardly bias pi1.

Will have to redo a lot of these analyses considering all test traits
