---
title: "IdentifyBrianaPacBioSamples"
output: html_document
date: '2023-05-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

Briana performed PacBio sequenced on 5 LCL lines Iwant to do more experiments to complement those datasets, but I'm not sure of what lines she used. I used QTLtools mbv to match her sewuencing data to the genotype data in 1000 genome's. Let's identify which samples she used here...


```{r}
library(tidyverse)
library(data.table)

dat <- Sys.glob("../code/QC/mbvLongReads/data/GM*.txt") %>%
  setNames(str_replace(., "../code/QC/mbvLongReads/data/(.+?).txt", "\\1")) %>%
  lapply(fread) %>%
  bind_rows(.id="Sample")


dat %>%
    ggplot(aes(x=perc_het_consistent, y=perc_hom_consistent,  label=SampleID)) +
    geom_text(size=1.5) +
    theme_minimal() +
    theme(legend.position="none") +
  facet_wrap(~Sample)


```

Looks like all the sample can be pretty unambigously identified. Let's print those sample names...


```{r}
dat %>%
  group_by(Sample) %>%
  filter(perc_het_consistent==max(perc_het_consistent)) %>%
  ungroup() %>%
  distinct(SampleID, Sample)

dat %>%
  group_by(Sample) %>%
  filter(perc_het_consistent==max(perc_het_consistent)) %>%
  ungroup() %>%
  distinct(SampleID)
```

Ok, so these are the samples to wake up:

NA18500
NA19153
NA18504
NA19144
NA19239

