---
title: "20240520_ScratchForEditorRequestedFigChanges"
author: "Benjamin Fair"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Introduction

This is your R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown, see [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/).

## R Code Chunk

```{r}
library(tidyverse)
```

## small molecule figs

```{r}
dat.sm <- read_tsv("../data/FigS23_Join_Fig5DE.dat.tsv.gz")

dat.sm %>%
    filter(!is.na(ColorEffect)) %>%
    distinct(gene, ColorEffect, .keep_all=T) %>%
    count(ColorEffect)
```

## Sample size of RNA-seq datasets

```{r}
samples <- read_tsv("../code/config/samples.tsv")

samples %>%
    filter(Assay == "RNA-seq") %>%
    distinct(Phenotype, IndID) %>%
    count(Phenotype)
```

## NUDT14 and D2GHGDH P, beta, q, PPFC

```{r}

```

### beta beta scatter exact P

### Intron len in risdi-exon containing introns, exact P