---
title: "202211107_Figures_sQTLeQTL_effects"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```


## Intro

Here I will plot some stuff related to prospective figure3 that links "unexplained" (read: "Non-enhancer/promoter" based) QTLs, largely to sQTL/NMD effects.

### set theme

```{r}
library(tidyverse)
library(RColorBrewer)
library(data.table)

# Set theme
theme_set(
  theme_classic() +
  theme(text=element_text(size=16,  family="Helvetica")))

#test theme
p <- ggplot(mtcars, aes(mpg, wt, color=as.factor(cyl))) +
  geom_point() +
  scale_colour_brewer(palette = "Set2", type="qual")
p
```

## Figure1

### highlight prevalence of NMD-targeting and unannotated introns in chRNA

### Numbers of QTLs

## Figure2

### heatmap of effects between types of phenotypes highlighting the expected strong enrichment in enhancers/promoters

### eQTLs colocalize with many different xQTLs, but chromatin effects are the strongest (something like the bar plot I showed in lab meeting with the number of eGenes that colocalize with an xQTL)

## Figure3

### Scatter plot of effect sizes for unannotated, basic, non-basic, and NMD-targeting introns. Estimate of number of primary (and maybe secondary) eQTLs explained by splicing mechanisms

```{r}

```

