---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Welcome to my research website.

This project will investigate the correlation of genetic effects on chromatin, splicing, transcription, and complex phenotypes, using naturally occuring human genetic variation.

We are still collecting data, but I have also been perusing around published data a bit.

Here are links to all of my rendered analysis Rmarkdowns (code [here](https://github.com/bfairkun/ChromatinSplicingQTLs/tree/master/analysis)):

```{r, results = 'asis'}
library(tidyverse)
rmds <- list.files(path = "../analysis/", pattern="^\\d+.+")
rmd_htmls <- str_replace(rmds, "Rmd$", "html")

for(i in rmd_htmls){
  cat("[", i, "](", i,")\n\n", sep="")
}
```

Also, a list of links to ipynb files that were incorporated into the snakemake. Since github renders ipynbs directly, this is not a relative link from the site folder, but rather links directly to the file to view on github (code [here](https://github.com/bfairkun/ChromatinSplicingQTLs/tree/master/code/notebooks)):

```{r, results = 'asis'}
ipynbs <- list.files(path = "../docs/", pattern="^\\d+.+ipynb$")

for(i in ipynbs){
  cat("[", i, "](","https://github.com/bfairkun/ChromatinSplicingQTLs/blob/master/docs/", i,")\n\n", sep="")
}

```

## Notebooks for figures
```{r, results = 'asis'}
library(tidyverse)
rmds <- c(list.files(path = "../analysis/", pattern="^Figures.+"), list.files(path = "../analysis/", pattern="^MakeFi.+"))
rmd_htmls <- str_replace(rmds, "Rmd$", "html")

for(i in rmd_htmls){
  cat("[", i, "](", i,")\n\n", sep="")
}
```

