---
title: "20240423_title"
author: "Benjamin Fair"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Introduction

This is your R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown, see [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/).

## R Code Chunk

```{r}
# source(file.path(Sys.getenv(if (.Platform$OS.type == "windows") "USERPROFILE" else "HOME"), ".vscode-R", "init.R"))
# if (interactive() && Sys.getenv("TERM_PROGRAM") == "vscode") {
#   if (requireNamespace("httpgd", quietly = TRUE)) {
#     options(vsc.plot = FALSE)
#     options(device = function(...) {
#       httpgd::hgd(silent = TRUE)
#       .vsc.browser(httpgd::hgd_url(history = FALSE), viewer = "Beside")
#     })
#   }
# }
# 
# options(vsc.plot = "Beside")
# options(vsc.dev.args = list(width = 800, height = 600))
# options(vsc.viewer = "Beside")

```

```{r}
library(edgeR)

data(mtcars)
plot(mtcars)

A <- c(1,2,3)
B <- c(2,3,4)

plot(A, B)

```

## More Information

For more information on customizing the behavior of the `rmarkdown` output format, see the [documentation](https://rmarkdown.rstudio.com/html_document_format.html).

For more details on using R Markdown, see [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/).
